The Shape Matching Element Method: Direct Animation of Curved Surface Models

### 2021-04-07 paper/sections/intro.tex
fixed typos

NURBS (Non-uniform Rational B Splines) are a popular higher-order modelling primitive which 
are used for computer-aided design (CAD), computational fabrication and computer animation. 
NURBS primitives were the first geometric representation used for physics-based animation~\cite{10.1145/37402.37427}, yet,
despite over three decades of research, animation of curved models such as those built with NURBS, remains a challenge.

>>>>

NURBS (Non-uniform Rational \revise{B-Splines}) are a popular higher-order modelling primitive which 
are used for computer-aided design (CAD), computational fabrication and computer animation. 
NURBS primitives were the first geometric representation used for physics-based animation~\cite{10.1145/37402.37427}, yet,
despite over three decades of research, animation of curved models such as those built with NURBS, remains a challenge.

### 2021-04-07 paper/sections/methods.tex
fixed typos

Our algorithm acts on objects composed of multiple NURBS surfaces~(\reffig{NURBS}).
\begin{figure}[h]
    \includegraphics[width=\columnwidth]{figures/nurbs_patch}
    \caption{A cubic NURBS patch with 16 control points.}
    \label{fig:NURBS}
\end{figure}
The three-dimensional position, $\defX$, of any point on a NURBS surface can be written as 
\begin{equation}
\label{eqn:nurbs_srf}
\defX\left(u,v\right) = \sum_{i=1}^n\sum_{j=1}^m \phi_{ij}\left(u,v\right)\vc{q}_{ij},
    %\mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    %{\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    %\text{,}
\end{equation}
where $u$ and $v$ $\in \real$ are the coordinates in the 2D parametric domain, $n$ and $m$ are the number of control points in the $u$ and $v$ directions, $\mathbf{q}_{ij}\in\real^3$ are the control points and $\phi_{ij}\left(u,v\right)\in\real$ are the NURBS
basis functions given by 
\begin{equation*}
    \phi_{ij}\left(u,v\right) = \frac{\omega_{i,j}\beta_{ik}(u)\beta_{jl}(v)}{\sum_{r=1}^{n}\sum_{s=1}^{m} \omega_{i,j} \beta_{ik}(u)\beta_{jl}(v)},
\end{equation*} where $\beta_{ik}$ (resp. $\beta_{jl}$) is the $i^{th}$ ($j^{th}$) B-spline basis of order $k$ ($l$) and $\omega_{i,j}$ is the NURBS weight.

@@

We can define a polynomial deformation from $\refX = \begin{pmatrix} X & Y & Z \end{pmatrix}^T$ to $\defX$  as 
\begin{equation}
    \defX\left(\refX\right) = 
    \underbrace{\begin{bmatrix}
    \mathbf{p}^T(\refX-\bar{\refX}) & \mathbf{0} & \mathbf{0} & 1 & 0 & 0\\
    \mathbf{0} & \mathbf{p}^T(\refX-\bar{\refX}) & \mathbf{0} & 0 & 1 & 0\\
    \mathbf{0} & \mathbf{0} & \mathbf{p}^T(\refX-\bar{\refX}) & 0 & 0 & 1\\
    \end{bmatrix}}_{\Pmatx}
    \Pcox,
    \label{eq:projection}
\end{equation} where $\mathbf{p}^T\left(\refX\right)$ is the polynomial basis vector \newline (e.g., $\begin{bmatrix} X & Y & Z & X\cdot Y & X\cdot Z & Y\cdot Z& X\cdot X & Y\cdot Y & Z\cdot Z \end{bmatrix}$ ),
$\bar{\refX}$ is an a priori chosen origin around which to deform, $\Pco$ are the coefficients for the 
non-constant part of the polynomial and $\tr$ is the translation of the deformation origin. 
We can compute the shape matching deformation by solving~\cite{10.1145/1073204.1073216}
\begin{equation}
   \Pcox^* = \arg\min_{\Pco,\tr} \int \left\lvert\left\lvert\Pmatx\Pcox - \Jnurbs\vc{q}\right\rvert\right\rvert^2_2 dudv,
\end{equation} where we have suppressed the dependence of $\refX$ and $\Jnurbs$ on $u$ and $v$ for the sake of brevity.

@@

\paragraph*{Relationship to the Virtual Element Method}
For a non-planar NURBS part, a sufficient number of quadrature points will make ~\refeq{shapematch} will be well-posed, allowing us to write: $\Pcox^* = \Pi\vc{q}$ where
\begin{equation*}
    \Pi = \left(\samp{\Pmat}^T\wt\samp{\Pmat}\right)^{-1}\samp{\Pmat}^T\wt\samp{\Jnurbs},
\end{equation*} is a projection from the space of control points onto the space of polynomials. 
This projection serves an identical purpose (and is mathematically equivalent up to choice of metric) to the projection operator used in VEM~\cite{10.1142/S021820251440003X}.
VEM uses a single polynomial projection per polyhedral element to solve differential equations on complex volumetric meshes.
We interpret shape matching as converting our NURBS part into a single Virtual Element suitable for elastodynamics simulation.
Unfortunately, a single polynomial deformation space will not be suitably expressive for even moderately complex models.
Previous work~\cite{10.1145/1073204.1073216,10.1145/1275808.1276480} used additional lattices or partitions of the reference space to 
increase the number of polynomials in the shape matching. In contrast, our meshless Shape Matching Element Method works directly on the NURBS
geometry itself.

@@

\dave{Ty: a bit more here, show the transformation of the integral via divergence theorem then show the final discrete version. Clearly define the quadrature rules.}
The key idea behind raycasting quadrature is to transform a domain integral of the form $I=\int_{\Omega} f(X,Y,Z) d\Omega$ to an easier form that enables us to select quadrature points in an entirely meshless manner. The first transformation is to convert the domain of the integral to an \textit{auxiliary domain}, $\Omega_R$ on which integration is easy to compute. In our case we take this domain to be an axis-aligned bounding box around our model, and rewrite integration as $I=\int_{\Omega_R} g(X,Y,Z) d\Omega_R$ where $g$ evaluates to zero outside of $\Omega$, but equals $f$ inside $\Omega$. Next we apply the Divergence theorem to rewrite our integral as
\newcommand{\Zeta}{\mathrm{Z}}
\begin{equation}
\int_{\Omega_R} g(X,Y,Z) d\Omega_R = \int_{\Gamma_R}\left(\int_{X_1}^{X_2} g(\alpha, Y, Z) d\alpha)\right)\mathbf{n} d\Gamma_R.
\end{equation}
where $\Gamma_R$ is the boundary of the bounding box with normals $\mathbf{n}$ and $X_1$ and $X_2$ represent the minimum and maximum values of the box along the $x$-axis. With this form, all we require to evaluate the outer integral is to integrate over the $y$-$z$ plane located at $X_1$. With this simple form, the normal term may be excluded as it always equals $1$ \dave{isn;t it a vector $(1,0,0)$ or something like that ? }in the case of an axis-aligned bounding box. Finally, raycasting is used to evaluate the line integral along the x-axis. From here, we find all the intersection intervals where the ray is inside $\Omega$, and select quadrature points uniformly along this interval. Thus, this transformation from the original volume integral into an integral over a $y$-$z$ plane plus axis-aligned rays enables us to integrate the volumes of our NURBS models in a completely meshless approach.

@@

\begin{equation}
    \defX\left(\refX\right) = \underbrace{\shapef\left(\refX\right)\bldmat\Pi}_{\Jkin\left(\refX\right)}\vc{q},
\end{equation} where $\shapef\in\real^{3\times n} = \begin{bmatrix}P_1\left(\refX\right) \dots P_m\left(\refX\right)\end{bmatrix}$, $m$ is the number of 
parts in the object, $\bldmat$ is the diagonal matrix of blending weights, $\Pi$ is the global projection operator and $\Jkin\left(\refX\right)$ is the kinematic Jacobian.

>>>>

Our algorithm acts on objects composed of multiple NURBS surfaces~(\reffig{NURBS}).
\begin{figure}[h]
    \includegraphics[width=\columnwidth]{figures/nurbs_patch}
    \caption{A cubic NURBS patch with 16 control points.}
    \label{fig:NURBS}
\end{figure}
The three-dimensional position, $\defX$, of any point on a NURBS surface can be written as 
\begin{equation}
\label{eqn:nurbs_srf}
\defX\left(u,v\right) = \sum_{i=1}^n\sum_{j=1}^m \phi_{ij}\left(u,v\right)\vc{q}_{ij},
    %\mathbf{x}(u,v) = \frac{\sum_{i=1}^{n}\sum_{j=1}^{m}  \mathbf{p}_{i,j} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    %{\sum_{i=1}^{n}\sum_{j=1}^{m} w_{i,j} B_{i,k}(u)B_{j,l}(v)}
    %\text{,}
\end{equation}
where $u$ and $v$ $\in \real$ are the coordinates in the 2D parametric domain, $n$ and $m$ are the number of control points in the $u$ and $v$ directions, $\mathbf{q}_{ij}\in\real^3$ are the control points and $\phi_{ij}\left(u,v\right)\in\real$ are the NURBS
basis functions given by 
\revise{
\begin{equation*}
    \phi_{ij}\left(u,v\right) = \frac{\omega_{i,j}\beta_{ik}(u)\beta_{jl}(v)}{\sum_{r=1}^{n}\sum_{s=1}^{m} \omega_{r,s} \beta_{rk}(u)\beta_{sl}(v)},
\end{equation*}} where $\beta_{ik}$ (resp. $\beta_{jl}$) is the $i^{th}$ ($j^{th}$) B-spline basis of order $k$ ($l$) and $\omega_{i,j}$ is the NURBS weight.

@@

We can define a polynomial deformation from $\refX = \begin{pmatrix} X & Y & Z \end{pmatrix}^T$ to $\defX$  as 
\revise{
\begin{equation}
    \defX\left(\refX\right) = 
    \underbrace{\begin{bmatrix}
    \mathbf{p}(\refX-\bar{\refX})^T & \mathbf{0} & \mathbf{0} & 1 & 0 & 0\\
    \mathbf{0} & \mathbf{p}(\refX-\bar{\refX})^T & \mathbf{0} & 0 & 1 & 0\\
    \mathbf{0} & \mathbf{0} & \mathbf{p}(\refX-\bar{\refX})^T & 0 & 0 & 1\\
    \end{bmatrix}}_{\Pmatx}
    \Pcox,
    \label{eq:projection}
\end{equation} where $\mathbf{p}\left(\refX\right)^T$} is the polynomial basis vector \newline (e.g., $\begin{bmatrix} X & Y & Z & X\cdot Y & X\cdot Z & Y\cdot Z& X\cdot X & Y\cdot Y & Z\cdot Z \end{bmatrix}$ ),
$\bar{\refX}$ is an a priori chosen origin around which to deform, $\Pco$ are the coefficients for the 
non-constant part of the polynomial and $\tr$ is the translation of the deformation origin. 
We can compute the shape matching deformation by solving~\cite{10.1145/1073204.1073216}
\begin{equation}
   \Pcox^* = \arg\min_{\Pco,\tr} \int \left\lvert\left\lvert\Pmatx\Pcox - \Jnurbs\vc{q}\right\rvert\right\rvert^2_2 dudv,
\end{equation} where we have suppressed the dependence of $\refX$ and $\Jnurbs$ on $u$ and $v$ for the sake of brevity.

@@

\paragraph*{Relationship to the Virtual Element Method}
For a non-planar NURBS part, a sufficient number of quadrature points \revise{will make ~\refeq{shapematch} well-posed}, allowing us to write: $\Pcox^* = \Pi\vc{q}$ where
\begin{equation*}
    \Pi = \left(\samp{\Pmat}^T\wt\samp{\Pmat}\right)^{-1}\samp{\Pmat}^T\wt\samp{\Jnurbs},
\end{equation*} is a projection from the space of control points onto the space of polynomials. 
This projection serves an identical purpose (and is mathematically equivalent up to choice of metric) to the projection operator used in VEM~\cite{10.1142/S021820251440003X}.
VEM uses a single polynomial projection per polyhedral element to solve differential equations on complex volumetric meshes.
We interpret shape matching as converting our NURBS part into a single Virtual Element suitable for elastodynamics simulation.
Unfortunately, a single polynomial deformation space will not be suitably expressive for even moderately complex models.
Previous work~\cite{10.1145/1073204.1073216,10.1145/1275808.1276480} used additional lattices or partitions of the reference space to 
increase the number of polynomials in the shape matching. In contrast, our meshless Shape Matching Element Method works directly on the NURBS
geometry itself.

@@

\dave{Ty: a bit more here, show the transformation of the integral via divergence theorem then show the final discrete version. Clearly define the quadrature rules.}
The key idea behind raycasting quadrature is to transform a domain integral of the form $I=\int_{\Omega} f(X,Y,Z) d\Omega$ to an easier form that enables us to select quadrature points in an entirely meshless manner. The first transformation is to convert the domain of the integral to an \textit{auxiliary domain}, $\Omega_R$ on which integration is easy to compute. In our case we take this domain to be an axis-aligned bounding box around our model, and rewrite integration as $I=\int_{\Omega_R} g(X,Y,Z) d\Omega_R$ where $g$ evaluates to zero outside of $\Omega$, but equals $f$ inside $\Omega$. Next we apply the Divergence theorem to rewrite our integral as
\newcommand{\Zeta}{\mathrm{Z}}

\begin{equation}
\int_{\Omega_R} g(X,Y,Z) d\Omega_R = \int_{\Gamma_R}\left(\int_{X_1}^{X_2} \revise{g(\alpha, Y, Z) d\alpha}\right)\mathbf{n} d\Gamma_R.
\end{equation}

where $\Gamma_R$ is the boundary of the bounding box with normals $\mathbf{n}$ and $X_1$ and $X_2$ represent the minimum and maximum values of the box along the $x$-axis. With this form, all we require to evaluate the outer integral is to integrate over the $y$-$z$ plane located at $X_1$. With this simple form, the normal term may be excluded as it always equals $1$ \dave{isn;t it a vector $(1,0,0)$ or something like that ? }in the case of an axis-aligned bounding box. Finally, raycasting is used to evaluate the line integral along the x-axis. From here, we find all the intersection intervals where the ray is inside $\Omega$, and select quadrature points uniformly along this interval. Thus, this transformation from the original volume integral into an integral over a $y$-$z$ plane plus axis-aligned rays enables us to integrate the volumes of our NURBS models in a completely meshless approach.

@@

\begin{equation}
    \defX\left(\refX\right) = \underbrace{\shapef\left(\refX\right)\bldmat\Pi}_{\Jkin\left(\refX\right)}\vc{q},
\end{equation} where $\shapef\in\real^{3\times n} = \begin{bmatrix}P_1\left(\refX\right) \dots P_m\left(\refX\right)\end{bmatrix}$, \revise{$n$ is of size $3(km + |\XBs|)$ with $k$ being the number of terms in each polynomial, $m$ is the number of parts, and $|\XBs|$ is the number of deformation origins. \ty{previously said:$m$ is the number of 
parts in the object}} $\bldmat$ is the diagonal matrix of blending weights, $\Pi$ is the global projection operator and $\Jkin\left(\refX\right)$ is the kinematic Jacobian.
